---
title: "Task 8"
output: 
  github_document: default
---
### The following code takes the values from "two" selected columns and adds it together into a new column and can be renamed as specified by the "user".
A dataset "test_pre" was tested for the function and generated a warning that states the data not in dataframe format. 

```{r}
library(dplyr)
library(lazyeval)
test_data <- read.csv("test_data.csv")
str(test_data)
col1_col2 <- function(df, list_of_cols, new_col) 
  { 
   if(is.data.frame(df)!= TRUE)
    warning("not a data frame")
  df %>% 
  mutate_(.dots = ~Reduce(`+`, .[list_of_cols])) %>% 
  setNames(c(names(df), new_col))
  }

head(col1_col2(test_data, c("Pre.Photosynthesis", "Post.Photosynthesis"), "Final_Photosynthesis"))
```

### The function uses a for loop and a package microbenchmark to calculate the sum of the elements "your_fun" vector, that returned 50005000 values.

#### For Loop
##### Takes the summary of all the squared values in loop_test_comparison

```{r}
library(microbenchmark)
# takes the summary of all the squared values in loop_test_comparison
loop_test_comparison <- c(1:10^4)
your_fun <- function (loop_test_comparison) 
  {for(val in loop_test_comparison)
  sums <- sum(loop_test_comparison) 
  sum
  print(sums)
  }
# takes the summary of all squared values for each rows because of "for loop"
your_fun(1:10^4)
test.vec <- 1:10^4
microbenchmark(your_fun(test.vec), sum(test.vec)) #your_fun takes longer because the for loop 
```
`
##### Takes the summary of all squared values for each rows because of "for loop"
```{r} 
your_fun(1:10^4)
test.vec <- 1:10^4
microbenchmark(your_fun(test.vec), sum(test.vec)) 
```
### In general, #your_fun takes longer because for the for loop option.
