---
title: "Task 7"
author: "Marife Anunciado"
date: "March 3, 2017"
output: html_document
---
```{r include=FALSE}
library(dplyr)
library(ggplot2)
library(nasaweather)
library(nycflights13)
library(babynames)
library(stringr)
library(reshape)
```
## Using the nycflights13::weather

Outliers in wind speed was rejected through a function **filter**, in this case wind_speed of over a value of 250 was discarded. 

```{r echo=FALSE, warning=FALSE}
flights_filter3 <- nycflights13::weather %>% 
  filter(wind_speed < 250) %>% group_by(wind_speed)
flights_filter3 %>% group_by(origin,wind_dir,wind_speed) %>%
      summarise(med_WS = median(wind_speed, na.rm=TRUE)) %>%
      mutate(med_WS) %>%
      select(origin,wind_dir,med_WS)
summary_median_WS <- flights_filter3 %>% group_by(origin,wind_dir,wind_speed) %>%
  summarise(med_WS = median(wind_speed, na.rm=TRUE)) %>%
  mutate(med_WS) %>%
  select(origin,wind_dir,med_WS)
summary_median_WS %>% ggplot(aes(x=med_WS, y=wind_dir, color=origin)) + 
  geom_point() + facet_grid(~ origin) +
  labs(x= "Wind speed, median", y= "Wind direction") + theme_classic() +
  ggsave("summary_median_WS_2.png", width = 5, height = 5)
```

Wind direction of **220 at EWR**; **200 at JFK** and **180 at LGA** airports has the highest median. Each airport's wind speed over wind direction was specifically shown as follows:

```{r echo=FALSE, warning=FALSE}
EWR_median_final <- summary_median_WS %>% 
  filter(origin == "EWR") %>% arrange(desc(wind_dir))
EWR_median_final %>% ggplot(aes(x=wind_dir,fill = med_WS)) + 
  geom_bar(width = 8, colour = "blue") + coord_polar(theta = "x") + 
  labs(title = "Median Wind Speed over Direction in EWR")

JFK_median_final <- summary_median_WS %>% 
  filter(origin == "JFK") %>% arrange(desc(wind_dir))

JFK_median_final %>% ggplot(aes(x=wind_dir,fill = med_WS)) + 
  geom_bar(width = 8, colour = "red") + coord_polar(theta = "x") + 
  labs(title = "Median Wind Speed over Direction in JFK")

LGA_median_final <- summary_median_WS %>% 
  filter(origin == "LGA") %>% arrange(desc(wind_dir))

LGA_median_final %>% ggplot(aes(x=wind_dir,fill = med_WS)) + 
  geom_bar(width = 8, colour = "purple") + coord_polar(theta = "x") + 
  labs(title = "Median Wind Speed over Direction in LGA")

```

## Using nycflights13::flights and nycflights13::airlines: airlines and its median distance (longest to shortest distance) travelled originating from JFK airport was displayed. 

```{r echo=FALSE, warning=FALSE}
combined_flightsairlines <- full_join(nycflights13::flights, nycflights13::airlines, by = "carrier") 
JFK_flightairlines <- combined_flightsairlines %>% filter(!grepl("EWR|LGA", origin)) 
JFK_median <- JFK_flightairlines %>% group_by(name) %>% 
  summarise(median= median(distance, na.rm=TRUE)) %>% arrange(desc(median)) 
JFK_median %>% ggplot(aes(x=median, y = name)) + 
  geom_col(width = 2, colour = "yellow") + theme_dark() 
```


## The plot below shows the wide-format data frame of number of flights originating from EWR airport at each month from each airlines.

```{r echo=FALSE, warning=FALSE, results="asis"}
EWR_flightairlines <- combined_flightsairlines %>% filter(!grepl("JFK|LGA", origin))
EWR_wide2 <- cast(EWR_flightairlines, name~month,value = "flight")
knitr::kable(EWR_wide2)
```

## The ten most common male and female names and their popularity from 1880 to 2014 were shown in the plots below:

```{r echo=FALSE, warning=FALSE}
babynames_2014 <- babynames %>% 
  filter(year > 2013) %>%
  arrange(desc(prop)) 
  
pop2014_names <- babynames %>%
  filter(name == "Noah" | 
           name == "Liam" | 
           name == "Mason" | 
           name =="Jacob"| 
           name =="William" | 
           name =="Ethan" | 
           name =="Michael" | 
           name == "Alexander" | 
           name =="James" | 
           name == "Daniel" | 
           name =="Emma" | 
           name =="Olivia" | 
           name == "Sophia" | 
           name =="Isabella" | 
           name =="Ava" | 
           name =="Mia" | name =="Emily" | 
           name =="Abigail" | 
           name =="Madison" | 
           name =="Charlotte")

#only includes 2014 most popular names
# male
pop2014_names_Male <- babynames %>%
  filter(name == "Noah" | 
           name == "Liam" | 
           name == "Mason" | 
           name =="Jacob"| 
           name =="William" | 
           name =="Ethan" | 
           name =="Michael" | 
           name == "Alexander" | 
           name =="James" | 
           name == "Daniel") 
           
pop2014_names_Male %>%
ggplot(aes(year,prop, color=name)) +
geom_point(size=1) + xlab("Year") + ylab("Proportion")
ggsave ("pop2014_names_Male.png", width = 10, height = 5)

# female
pop2014_names_female <- babynames %>%
  filter( name =="Emma" | 
           name =="Olivia" | 
           name == "Sophia" | 
           name =="Isabella" | 
           name =="Ava" | 
           name =="Mia" | name =="Emily" | 
           name =="Abigail" | 
           name =="Madison" | 
           name =="Charlotte")

pop2014_names_female %>%
  ggplot(aes(year,prop, color=name)) +
  geom_point(size=1) + xlab("Year") + ylab("Proportion")
ggsave ("pop2014_names_female.png", width = 10, height = 5)
```

The most popular names in 1896, 1942 and 2014 was shown below as well as their proportion.

```{r echo=FALSE, warning=FALSE, results="asis"}
baby_1896 <- babynames %>% 
  filter(year == 1896) %>%
  arrange(desc(prop))     
female_1896 <- baby_1896 %>% 
  filter(!grepl("M", sex)) %>% arrange(desc(prop))  
female_1896_pop <- female_1896 %>%
  filter(name == "Martha" | name == "Esther" | name == "Frances" | name =="Edith") 
baby_1942 <- babynames %>% 
  filter(year == 1942) %>%
  arrange(desc(prop))     
female_1942 <- baby_1942 %>% 
  filter(!grepl("M", sex)) %>% arrange(desc(prop))  
female_1942_pop <- female_1942 %>%
  filter(name == "Marilyn" | name == "Diane" | name == "Martha" | name =="Frances") 
baby_2014 <- babynames %>% 
  filter(year == 2014) %>%
  arrange(desc(prop))      
female_2014 <- baby_2014 %>% 
  filter(!grepl("M", sex)) %>% arrange(desc(prop))  
female_2014_pop <- female_2014 %>%
  filter(name == "Brooklyn" | name == "Lily" | name == "Hannah" | name =="Layla") 
common_names <- bind_rows(list(female_1896_pop, female_1942_pop, female_2014_pop)) 
knitr::kable(common_names)
```

## Using nasaweather datasets on storms
  1. calculate the average disturbance for each year
  2. plot the disturbance from year 1996 to 2000
  3. look at the weather disturbance that occurred in 1995
  4. determine the months with weather disturbances

```{r echo=FALSE, warning=FALSE}
weather_2 <- nasaweather::storms %>% group_by(name, year, type) %>% 
  summarise(Disturbance =max(seasday, na.rm=TRUE)) %>%
  mutate(Disturbance) %>% select(name, year, type, Disturbance)

weather_2 %>% ggplot(aes(x=year, y=Disturbance, color=type)) + 
  geom_point(position = "jitter") + facet_grid(~ type)  + 
  xlab("Year") + ylab("Weather Disturbance, mean") + theme_dark()
ggsave("weather1.png", width = 10, height = 5)

weather_2 %>% ggplot(aes(x= year, y=Disturbance, color=type)) + 
  geom_jitter(stat="identity")  + 
  xlab("Year") + ylab("Weather Disturbance, mean") + theme_classic()
ggsave("weather2.png", width = 8, height = 5)

Weather_1995 <- weather_2 %>% filter(year == 1995) %>% group_by(type)
Weather_1995 %>% ggplot(aes(x=name, y=Disturbance, color=type)) + 
  geom_point(size=2) + xlab("Name") + ylab("Weather Disturbance, mean") + theme_grey()
ggsave("weather3.png", width = 12, height = 5)
```
